Attribute fusion in a latent process model for time series of graphs

### 2011-04-29 ssp2011.tex
Revised some proofs

concerned with the extension of the results in

@@

graphs. We state here some the relevant results. Sketches of the
proofs of these results will be provided in
\S~\ref{sec:proofs-some-stated}. \\ \\

@@

@@

@@

We consider first the case where $t < t^{*}$. Then, conditional on
$|N(v)| = l$, we have

@@

Therefore, for sufficiently large $l$, by the multivariate central
limit theorem, we have

@@

  \Bigl(\psi_\lambda(v;t) \, \lvert \, |N(v)| = l \Bigr) \sim
  \mathcal{N}\Bigl(\tbinom{l}{2} \langle \lambda, \pi_{00} \rangle,
  \tbinom{l}{2} \langle \lambda, \eta_{00} \lambda \rangle \Bigr)

@@

Let $\lambda^{(2)}$ be the vector whose entries are element-wise
square of $\lambda$. Then it can be checked that 

@@

 \langle \lambda,
\eta_{00} \lambda \rangle = \langle \lambda^{(2)}, \pi_{00} \rangle -
  (\langle \lambda, \pi_{00} \rangle)^2

@@

@@

 \Bigl(\psi_\lambda(v;t) \,\lvert\,\, |N(v)| = l \Bigr) \sim C *  
 \mathrm{Bin}\Bigl(\tbinom{l}{2}, p\Bigr)

@@

  Let $\Psi_{\lambda}(t)$ be the (fused) scan statistic. Let
   \mu_0 &= (n-1) \langle \pi_{00}, 1 \rangle \\
 \sigma_0 &= \sqrt{(n-1) (\langle \pi_{00}, 1 \rangle - \langle
   \pi_{00}, 1 \rangle^2)} \\ 
 z_n &= \sqrt{2
      \log{n}}\Bigr) \\
   a_n &= N_{0} \frac{\langle \lambda, \pi_{00} \rangle}{\langle 1,
     \pi_{00} \rangle} + p\tbinom{N_{0}}{2} \\ 
   b_n &= (1 - \tfrac{p}{2} + p N_{0}) \frac{\sigma_0}{\sqrt{2 \log
  \Psi_{\lambda}(t) \rightsquigarrow \mathcal{G}(C a_n, C b_n)

@@

$\kappa(n,n_1,p_1,p_2,p_3)$. Let us define $B$, $C$, $E$ and $F$ by

@@

  B \sim \mathrm{Bin}(m, \langle 1, \pi_{00} \rangle); \quad C \sim \mathrm{Bin}(n-m-1,
 \langle 1, \pi_{01} \rangle) \\
  E \sim \mathrm{Bin}(m-1, \langle 1, \pi_{11} \rangle); \quad F \sim \mathrm{Bin}(n-m,
  \langle 1, \pi_{01} \rangle)

@@

where $z_m$ is
  z_m = \sqrt{2 \log m}\Bigl(1 - \frac{\log{\log{m}} + \log{4
      \pi}}{\log{m}}\Bigr)
The following result is adapted from \cite{rukhin:_limit_distr_graph_scan_statis}
    a_{n,m} &= N_{\kappa} \frac{\langle x, \pi_{01} \rangle}{\langle 1, \pi_{01} \rangle} \\ &+ C_1 p_1 \binom{N_\kappa}{2} + (C_2 p_2 -
    C_1 p_1)\binom{\mu_E}{2} \\ &+ (C_3 p_3 - C_1 p_1)\mu_E \mu_F 
    b_{n,m} = (1 - \tfrac{C_1 p_1}{2} + C_1 p_1
    \Psi_{\lambda}(t^{*}) \rightsquigarrow \mathcal{G}(a_{n,m}, b_{n,m})
  with $n-1$ trials and probability vector $\pi_0$. The following
  statements are made in the limit as $n \rightarrow \infty$ for fixed
  $K$. By the multivariate
  central limit theorem, we have
    \bm{\Gamma}(v;t) \sim \mathcal{N}((n-1)\pi_0,
  and therefore
    \label{eq:12}
    d_{\lambda}(v;t) \sim \mathcal{N}((n-1) \langle \lambda, \pi_0
  where $\eta_{00} = \mathrm{diag}(\pi_0) - \pi_0
  \pi_0^{T}$. $\Delta_{\lambda}(t)$ for $t < t^{*}$ is thus the
  maximum of a sequence of {\em dependent} Gaussian random
  ignored. $\Delta_{\lambda}(t)$ is then equivalent to the maximum of
  a sequence of {\em independent} $\mathcal{N}((n-1) \langle \lambda,

@@

@@

>>>>

concerned with adpating the results in

@@

graphs. We state here only the relevant results. The proofs are
delegated to the appendices. \\ \\

@@

The limiting distribution for $\Psi_{\lambda}(t)$ is Gumbel for both
the case where $t < t^{*}$ and for $t = t^{*}$. To aid the exposition,
we first introduce some notations. Let $\lambda^{(2)}$ be the
element-wise square of $\lambda$. Define $C_{00}$, $C_{01}$, $C_{11}$
and $p_{00}$, $p_{01}$, $p_{11}$ to be
  \label{eq:19}
  C_{00} = \tfrac{\langle \lambda^{(2)}, \pi_{00} \rangle}{\langle \lambda,
    \pi_{00}\rangle}; \quad p_{00} = \tfrac{(\langle \lambda, \pi_{00}
    \rangle)^{2}}{\langle \lambda^{(2)},\, \pi_{00} \rangle} \\
C_{11} = \tfrac{\langle \lambda^{(2)}, \pi_{11} \rangle}{\langle \lambda,
    \pi_{11}\rangle}; \quad p_{11} = \tfrac{(\langle \lambda, \pi_{11}
    \rangle)^{2}}{\langle \lambda^{(2)},\, \pi_{11} \rangle} \\
C_{10} = \tfrac{\langle \lambda^{(2)}, \pi_{11} \rangle}{\langle \lambda,
    \pi_{10}\rangle}; \quad p_{10} = \tfrac{(\langle \lambda, \pi_{10}
    \rangle)^{2}}{\langle \lambda^{(2)},\, \pi_{10} \rangle} 
Also let $z_n$ be defined, for any $n \in \mathbb{N}$, by

@@

  \label{eq:20}
   z_n = \sqrt{2
      \log{n}}\Bigr)

@@

  Let $\Psi_{\lambda}(t)$ be the scan statistic for $t < t^{*}$. Let
   \mu_0 &= (n-1) \langle 1, \pi_{00} \rangle \\
   \sigma_0 &= \sqrt{(n-1) (\langle 1, \pi_{00} \rangle - \langle
     1, \pi_{00} \rangle^2)} \\ 
    a_n &= N_{0} \frac{\langle \lambda, \pi_{00} \rangle}{\langle
      1, \pi_{00} \rangle} + C_{00}p_{00}\tbinom{N_{0}}{2} \\ 
   b_n &= (1 - \tfrac{C_{00}p_{00}}{2} + C_{00}p_{00} N_{0}) \frac{\sigma_0}{\sqrt{2 \log
  \frac{\Psi_{\lambda}(t) - a_{n}}{b_n} \rightsquigarrow
  \mathcal{G}(0, 1)
\end{theorem}%
We now move to the limiting distribution for $\Psi_{\lambda}(t^{*})$. 
Let us define $E$ and $F$ by
 %  B \sim \mathrm{Bin}(m, \langle 1, \pi_{00} \rangle); \quad C \sim \mathrm{Bin}(n-m-1,
 % \langle 1, \pi_{01} \rangle) \\
  E \sim \mathrm{Bin}(m-1, \langle \pi_{11}, 1 \rangle) \\ F \sim \mathrm{Bin}(n-m,
  \langle \pi_{01}, 1 \rangle)
\end{gather*}%
    a_{n,m} &=%  N_{\kappa} \frac{\langle x, \pi_{01} \rangle}{\langle
 %      1, \pi_{01} \rangle} \\
 % & +
    C_{00} p_{00} \Bigl(\tbinom{N_\kappa}{2} - \tbinom{\mu_E}{2} - \mu_E
    \mu_F\Bigr) \\ 
     & + C_{11} p_{11} \tbinom{\mu_E}{2} + C_{01} p_{01} \mu_E \mu_F 
    b_{n,m} = (1 - \tfrac{C_{00} p_{00}}{2} + C_{00} p_{00}
    \frac{\Psi_{\lambda}(t^{*}) - a_{n,m}}{b_{n,m}}  \rightsquigarrow \mathcal{G}(0,1)  \end{equation}

@@

@@

@@

@@

@@

  with $n-1$ trials and probability vector $\pi_0$. By the multivariate
  normal approximation to the multinomial, we have
    \bm{\Gamma}(v;t) \approx \mathcal{N}((n-1)\pi_0,
  and thus, for sufficiently large $n$ we have
    d_{\lambda}(v;t) = (1 + o_P(1)) \mathcal{N}((n-1) \langle \lambda, \pi_0
  where $\eta_{00} = \mathrm{diag}(\pi_0) - \pi_0 \pi_0^{T}$ and
  $o_P(1)$ denotes a random variables that converges to $0$ in
  probability. We can thus take $\Delta_{\lambda}(t)$ for $t < t^{*}$
  as the maximum of a sequence of {\em dependent} Gaussian random
  ignored. $\Delta_{\lambda}(t)$ can thus be taken as the maximum of
  {\em independent} $\mathcal{N}((n-1) \langle \lambda,

@@

@@

@@

\begin{proof}[Theorem~\ref{thm:2}]
The following statements are made conditional on $|N(v)| = l$. First
of all, we have

@@

@@

For large $l$, by the multivariate normal
approximation to the multinomial we have

@@

  \psi_\lambda(v;t) & = (1 + o_{P}(1)) \mathcal{N}\Bigl(\tbinom{l}{2} \langle \lambda, \pi_{00} \rangle,
  \tbinom{l}{2} \langle \lambda, \eta_{00} \lambda \rangle \Bigr) \\ 
  &= (1 + o_{P}(1)) C_{00} \mathcal{N}(\tbinom{l}{2} p_{00}, \tbinom{l}{2} p_{00}
  (1 - p_{00}))

@@

Now $p_{00} \leq 1$ with strict inequality whenever $\eta_{00}$ is
positive definite. We thus have, for sufficiently large $l$,
 \psi_\lambda(v;t) = (1 + o_P(1)) C_{00} *  
 \mathrm{Bin}\Bigl(\tbinom{l}{2}, p_{00}\Bigr)
Eq.~\eqref{eq:17} states that the locality statistics for the
attributed random graphs model under considation with $t < t^{*}$ can
be approximated by the locality statistics for an Erd\"{o}s-Renyi
graph with edge probability $p_{00}$. The claim then follows from
\cite[Theorem 1.1]{rukhin:_limit_distr_graph_scan_statis}.
